Hierarchically Supervised Latent Dirichlet Allocation

### 2011-10-27 paper/experiments.tex
Camera ready draft 2

Our dataset was gathered from the clinical data warehouse of
a large metropolitan hospital. It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. All included discharge summaries had associated ICD-9 Codes.
Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

@@

Our dataset contains 15,130 product descriptions for training and 1,000
for testing. The product descriptions are shorter than the discharge summaries
(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique codes.
Products are assigned on average 9.01 codes (std dev=4.91). The vocabulary
consists of the most frequent 30,000 words omitting stopwords.

@@

We evaluated HSLDA along with two closely related models against
the two datasets. The comparison models included sLDA with independent
regressors (hierarchical constraints on labels ignored), HSLDA fit by first
performing LDA then fitting tree-conditional regressions. These models were
chosen to  highlight several aspects of HSLDA including performance in the
absence  of hierarchical constraints, the effect of the combined inference, and
regression performance attributable solely to the hierarchical constraints.

@@

The gold standard for comparison was derived from the testing set in each
dataset. To make the comparison as fair as possible among models, ancestors of
observed nodes in the label hierarchy were ignored, observed nodes were
considered positive and descendents of observed nodes were considered to be
negative. In particular, since the sLDA model does not enforce the hierarchical
constraints, we establish a more equal footing by
considering only the observed labels as being positive, despite the fact
that, following the hierarchical constraints, ancestors must also be positive.
Such a gold standard will likely inflate the number of false positives because
the labels applied to any particular document are usually not as complete as
they could be.  ICD-9 codes, for instance, lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as
false positives~\citep{Birmetal2005}. 
However, given that the label space is often large (as in our examples) it is a
moderate assumption that erroneous false positives should not skew results
significantly.

>>>>

Our dataset was gathered from the NewYork-Presbyterian Hospital clinical data warehouse. 
It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. All included discharge summaries had associated ICD-9 Codes.
Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

@@

Our dataset contains 15,130 product descriptions for training and 1,000
for testing. The product descriptions are shorter than the discharge summaries
(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique categories.
Products are assigned on average 9.01 categories (std dev=4.91). The vocabulary
consists of the most frequent 30,000 words omitting stopwords.

@@

We evaluated HSLDA along with two closely related models against
the two datasets. The comparison models included sLDA with independent
regressors (hierarchical constraints on labels ignored) and HSLDA fit by first
performing LDA then fitting tree-conditional regressions. These models were
chosen to  highlight several aspects of HSLDA including performance in the
absence  of hierarchical constraints, the effect of the combined inference, and
regression performance attributable solely to the hierarchical constraints.

@@

The gold standard for comparison was derived from the testing set in each
dataset. To make the comparison as fair as possible among models, ancestors of
observed nodes in the label hierarchy were ignored, observed nodes were
considered positive and descendents of observed nodes were considered to be
negative. Note that this is different from our treatment of the observations during inference.
Since the sLDA model does not enforce the hierarchical
constraints, we establish a more equal footing by
considering only the observed labels as being positive, despite the fact
that, following the hierarchical constraints, ancestors must also be positive.
Such a gold standard will likely inflate the number of false positives because
the labels applied to any particular document are usually not as complete as
they could be.  ICD-9 codes, for instance, lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as
false positives~\citep{Birmetal2005}.
However, given that the label space is often large (as in our examples) it is a
moderate assumption that erroneous false positives should not skew results
significantly.

### 2011-10-27 paper/introduction.tex
Camera ready draft 2

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper we focus on
unstructured text data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes.
In this work we show how to combine 
these two sources of information using a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. The models and techniques that we develop in
this paper are applicable to other data as well, namely, any unstructured
representations of data that have been hierarchically classified (e.g.,~image
catalogs with bag-of-feature image representations).

There are several challenges entailed in incorporating a hierarchy of labels
into the model. Given a large set of potential labels (often thousands), each
instance has only a small number of labels associated to it. There are no
naturally occurring negative labeling in the data, and the absence of a label
cannot always be interpreted as a negative labeling.

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. For the sake of
simplicity, we focus on is-a hierarchies, but the model can be
applied to other hierarchies. We extend supervised latent Dirichlet
allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
supervision. We propose an efficient way to incorporate hierarchical
information into the model. We hypothesize that the context of labels within
the hierarchy provides valuable information about labeling.

We demonstrate our model on large, real-world datasets in the clinical and web
retail domains. We observe that hierarchical information is valuable when
incorporated into the learning and improves our primary goal of multi-label
classification. Our results show that a joint, hierarchical model outperforms a
classification with unstructured labels as well as a disjoint model, where the
topic modeling and the hierarchical classification are carried out
independently of each other.

>>>>

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper we focus on
unstructured text data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes.
In this work we show how to combine 
these two sources of information using a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. The models and techniques that we develop in
this paper are applicable to other data as well, namely, any unstructured
representations of data that have been hierarchically classified (e.g.,~image
catalogs with bag-of-feature representations).

There are several challenges entailed in incorporating a hierarchy of labels
into the model. Among them, given a large set of potential labels (often thousands), each
instance has only a small number of labels associated to it. Furthermore, there are no
naturally occurring negative labeling in the data, and the absence of a label
cannot always be interpreted as a negative labeling.

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. For the sake of
simplicity, we focus on ``is-a'' hierarchies, but the model can be
applied to other structured label spaces. We extend supervised latent Dirichlet
allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
supervision. We propose an efficient way to incorporate hierarchical
information into the model. We hypothesize that the context of labels within
the hierarchy provides valuable information about labeling.

We demonstrate our model on large, real-world datasets in the clinical and web
retail domains. We observe that hierarchical information is valuable when
incorporated into the learning and improves our primary goal of multi-label
classification. Our results show that a joint, hierarchical model outperforms a
classification with unstructured labels as well as a disjoint model, where the
topic model and the hierarchical classification are inferred
independently of each other.

### 2011-10-27 paper/model.tex
Camera ready draft 2

HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We
will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d}
\in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
= \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document
$d$.  Let there be $D$ such documents and let the size of the vocabulary be
$V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label labels
$l \in \mathcal{L}$, except root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
also in the set of labels. 
 We will for exposition purposes assume that this label set has hard ``is-a''
 parent-child constraints (explained later), although this assumption can be
 relaxed at the cost of more computationally complex inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we consider, only positive label applications are observed.

>>>>

HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We
will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d}
\in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
= \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document
$d$.  Let there be $D$ such documents and let the size of the vocabulary be
$V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label
$l \in \mathcal{L}$, except the root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
also in the set of labels. 
 We will for exposition purposes assume that this label set has hard ``is-a''
 parent-child constraints (explained later), although this assumption can be
 relaxed at the cost of more computationally complex inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we consider, only positive labels are observed.

### 2011-10-27 paper/related_work.tex
Camera ready draft 2

Other models that incorporate LDA and supervision include
LabeledLDA\citep{Ramage2009} and DiscLDA\citep{DiscLDA}.  Various applications of these models to 
computer vision and document networks have been explored
\citep{wangbleifeifei08,RelationalLDA} .
None of these models, however, leverage dependency structure in the label space.

>>>>

Other models that incorporate LDA and supervision include
LabeledLDA~\citep{Ramage2009} and DiscLDA~\citep{DiscLDA}.  Various applications of these models to 
computer vision and document networks have been explored~\citep{wangbleifeifei08,RelationalLDA} .
None of these models, however, leverage dependency structure in the label space.
